---
title: "ADAFinalProject"
author: "Mara McKown"
date: '2022-11-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ADA Final Project
### Install packages and libraries
```{r, message=FALSE}
pacman::p_load(readr, tidyverse, table1, dplyr, readxl, naniar, writexl, MASS, lmtest, stargazer, sandwich, broom)
```

### Upload Data
```{r}
df <- read_excel("/Users/maramckown/Documents/WashU/Classes/Fall 2022 Classes/Advanced Data Analysis/Final Project/ExploreMOhealth.data.xlsx", sheet = 3)
```

### Rename columns
```{r}
dput(names(df))
df <- df %>% 
  rename("zipcode" = "ZIP Code",
         "zipname" = "ZIP Name",
         "rankingindex" = "ZIP Health Ranking Index Score",
         "healthrank" = "ZIP Health Rank",
         "combinedindex" = "Health Factors Combined Index Score",
         "combinedrank" = "Health Factors Combined Rank",
         "teenpregrate" = "Teen Pregnancy per 1,000",
         "teenpregz" = "Teen Pregnancy per 1,000 Z-Score",
         "teenpregrank" = "Teen Pregnancy Rank (of 937)",
         "lbwrate" = "Low Birthweight per 1,000",
         "lbwz" = "Low Birthweight per 1,000 Z-Score",
         "lbwrank" = "Low Birthweight Rank (of 937)",
         "hsedupercent" = "High School Education (%)",
         "medianincome" = "Median Household Income ($)",
         "qualityliferank" = "Quality of Life Rank (of 937)",
         "withouths" = "Low Education (%)"
         )
```

### Recoding
```{r}
#make variable that is number of people with hs edu, by subtracting those without hs from 100
df$withhs <- 100 - df$withouths

# make high and low teen pregnancy cut off
df <- df %>% 
  mutate(teenpregh_l = case_when(teenpregrank >= 469 ~ "High",
                                 teenpregrank <= 468 ~ "Low"))
df$teenpregh_l <- factor(df$teenpregh_l, labels = c("High Teen Pregnancy Rates", "Low Teen Pregnancy Rates"))
table1(~teenpregh_l, data = df) 
table(df$teenpregrank, df$teenpregh_l)#just make sure it makes sense

table1(~ lbwrate + withouths + withhs, data = df) 

#change lbwrate and with hs to be whole numbers so can do Poisson (as they have two decimal places will multiply by 100 and the rate will change from out of 1,000 to out of 100,000)
df$lbw100k <- df$lbwrate * 100
df$withhs100k <- df$withhs * 100
```

### Descriptive Stats
```{r}
#add labels to variables
label(df$withhs100k)<-"High School Education (out of 100k)"
label(df$qualityliferank)<-"Quality of Life Ranking (of 937)"
label(df$medianincome)<-"Median Income ($)"
label(df$teenpregh_l)<-"Teen Pregnancy (High/Low)"
label(df$lbw100k)<-"Low Birth Weight (out of 100k)"

#descriptive table of outcomes separated by high vs. low teen pregnancy rates
table1(~ withhs100k + lbw100k | teenpregh_l, data = df)

#descriptive table of confounders
table1(~qualityliferank + medianincome, data = df)
table1(~qualityliferank + medianincome | teenpregh_l, data = df)

#full descriptive table
table1(~withhs100k + lbw100k + qualityliferank + medianincome | teenpregh_l, overall="Total", footnote='SD = standard deviation; teen pregnancy rates were categorized using the median zip code ranking of all zip codes in the state of Missouri as the cutoff value', data = df)

#graphs of outcome variables
lbwgraph <- ggplot(data = df, aes(lbw100k)) +
  geom_histogram(aes(y = ..density..), bins = 15) +
  geom_density() +
  labs(x = "Low Birth Weight (per 100,000 births)") +
  theme_bw()
ggsave(filename = "lbw.png", plot = lbwgraph,
       width = 7, height = 4, bg="white") 
hsgraph <- ggplot(data = df, aes(withhs100k)) +
  geom_histogram(aes(y = ..density..), bins = 15) +
  geom_density() +
  labs(x = "High School Education (per 100,000 people)") +
  theme_bw()
ggsave(filename = "hs.png", plot = hsgraph,
       width = 7, height = 4, bg="white") 
```

### Model looking at how high vs. low teen pregnancy rate impacts lbw (w/o confounding)
```{r}
#Models without confounding
##Poisson
model0 <- glm(lbw100k ~ teenpregh_l, family ="poisson", data = df)
summary(model0)
broom::tidy(model0, conf.int = TRUE, exponentiate = TRUE)

##Neg Binomial
model0nb <- glm.nb(lbw100k ~ teenpregh_l, control=glm.control(maxit=50),  data = df)
summary(model0nb)
broom::tidy(model0nb, conf.int = TRUE, exponentiate = TRUE)
#models had p<0.05 so decided to include confounding

#compare SEs
stargazer(model0, model0nb, title="Model Comparison",
          type="text",align=TRUE,single.row=TRUE, digits=6)

#robust SE and rest of model using robust SEs
robust0 <- tidy(coeftest(model0nb, vcov. = vcovHC(model0nb, type = 'HC3'))) 
robust0$IRR<- exp(robust0$estimate) #incidence rate ratio
robust0$lowCI <- exp(robust0$estimate - 1.96*robust0$std.error)
robust0$highCI <- exp(robust0$estimate + 1.96*robust0$std.error)
robust0
```

### Model looking at how high vs. low teen pregnancy rate impacts lbw (w/ confounding)
```{r}
#Models with confounding
##Poisson model
model1 <- glm(lbw100k ~ teenpregh_l + qualityliferank + medianincome, family ="poisson", data = df)
summary(model1)
broom::tidy(model1, conf.int = TRUE, exponentiate = TRUE)

##Negative Binomial model
model1nb <- glm.nb(lbw100k ~ teenpregh_l + qualityliferank + medianincome, control=glm.control(maxit=50),  data = df)
summary(model1nb)
broom::tidy(model1nb, conf.int = TRUE, exponentiate = TRUE)

#run lrtest to compare models
lrtest(model1, model1nb)
#models are significantly different, so overdispersion and should use negative binomial model

#compare SEs
stargazer(model1, model1nb, title="Model Comparison",
          type="text",align=TRUE,single.row=TRUE, digits=6)

#robust SE and rest of model using robust SEs
robust1 <- tidy(coeftest(model1nb, vcov. = vcovHC(model1nb, type = 'HC3'))) 
robust1$IRR<- exp(robust1$estimate) #incidence rate ratio
robust1$lowCI <- exp(robust1$estimate - 1.96*robust1$std.error)
robust1$highCI <- exp(robust1$estimate + 1.96*robust1$std.error)
robust1
```

### Model looking at how high vs. low teen pregnancy rate impacts high school education (w/o confounding)
```{r}
##No confounding
#Poisson model
model4 <- glm(withhs100k ~ teenpregh_l, family ="poisson", data = df)
summary(model2)
broom::tidy(model2, conf.int = TRUE, exponentiate = TRUE)

#Negative Binomial model
model4nb <- glm.nb(withhs100k ~ teenpregh_l, control=glm.control(maxit=50),  data = df)
summary(model2nb)
broom::tidy(model2nb, conf.int = TRUE, exponentiate = TRUE)

#run lrtest to compare models
lrtest(model4, model4nb)
#models are significantly different, so over dispersion and should use negative binomial model

#compare SEs
stargazer(model4, model4nb, title="Model Comparison",
          type="text",align=TRUE,single.row=TRUE, digits=6)

#robust SE and rest of model using robust SEs
robust4 <-tidy(coeftest(model4nb, vcov. = vcovHC(model4nb, type = 'HC3'))) 
robust4$IRR<- exp(robust4$estimate) #incidence rate ratio
robust4$lowCI <- exp(robust4$estimate - 1.96*robust4$std.error)
robust4$highCI <- exp(robust4$estimate + 1.96*robust4$std.error)
robust4
```


### Model looking at how high vs. low teen pregnancy rate impacts high school education (w/ confounding)
```{r}
##with confounding
#Poisson model
model2 <- glm(withhs100k ~ teenpregh_l + qualityliferank + medianincome, family ="poisson", data = df)
summary(model2)
broom::tidy(model2, conf.int = TRUE, exponentiate = TRUE)

#Negative Binomial model
model2nb <- glm.nb(withhs100k ~ teenpregh_l + qualityliferank + medianincome, control=glm.control(maxit=50),  data = df)
summary(model2nb)
broom::tidy(model2nb, conf.int = TRUE, exponentiate = TRUE)

#run lrtest to compare models
lrtest(model2, model2nb)
#models are significantly different, so over dispersion and should use negative binomial model

#compare SEs
stargazer(model2, model2nb, title="Model Comparison",
          type="text",align=TRUE,single.row=TRUE, digits=6)

#robust SE and rest of model using robust SEs
robust2 <-tidy(coeftest(model2nb, vcov. = vcovHC(model2nb, type = 'HC3'))) 
robust2$IRR<- exp(robust2$estimate) #incidence rate ratio
robust2$lowCI <- exp(robust2$estimate - 1.96*robust2$std.error)
robust2$highCI <- exp(robust2$estimate + 1.96*robust2$std.error)
robust2
```

#need to look to see if quality of life is an effect modifier for relationship between teen pregnancy and low birth weight
```{r}
#Negative Binomial model without interaction
model3nb <- glm.nb(lbw100k ~ teenpregh_l + qualityliferank + medianincome, control=glm.control(maxit=50),  data = df)
summary(model3nb)
broom::tidy(model3nb, conf.int = TRUE, exponentiate = TRUE)

#Negative Binomial model with interaction
model3nb_int <- glm.nb(lbw100k ~ teenpregh_l + qualityliferank + medianincome + qualityliferank*teenpregh_l, control=glm.control(maxit=50),  data = df)
summary(model3nb_int)
broom::tidy(model3nb, conf.int = TRUE, exponentiate = TRUE)

#run lrtest to compare models
lrtest(model3nb, model3nb_int)
#models are significantly different, so over dispersion and should use negative binomial model

#compare SEs
stargazer(model3, model3nb, title="Model Comparison",
          type="text",align=TRUE,single.row=TRUE, digits=6)

#robust SE and rest of model using robust SEs
robust3 <-broom::tidy(coeftest(model3nb, vcov. = vcovHC(model3nb, type = 'HC3'))) 
robust3$IRR<- exp(robust3$estimate) #incidence rate ratio
robust3$lowCI <- exp(robust3$estimate - 1.96*robust3$std.error)
robust3$highCI <- exp(robust3$estimate + 1.96*robust3$std.error)
robust3
```
